---
title: "01-exercises"
author: "Christopher Brown"
editor: "Jarus Singh"
date: "April 6, 2016"
output: html_document
---


## Exercise 1:

Write a function (f) that takes a vector of numbers, **x**, and returns a vector of numbers such that each element containing the product of every element of x except the element of x with the same index. 


### Example 

    > x <- c( 1, 5, 2, 8 ) 
    > f(x) 
    [1] 80 16 40 10
    # 5*2*8, 1*2*8, 1*5*8, 1*2*5  


### Solution 

```{r}
# Elegant solution
f <- function(x) { prod(x) / ifelse(x==0,1,x) }
```

```{r}
# Inelegant solution that I wrote during class
f_inelegant <- function(x) {
  # Output will be stored in y, initialize as empty vector
  y <- c()
  
  # Computes product of all numbers and store in product_all
  # Start loop
  for (i in 1:length(x)) {
    # If product_all doesn't exist, make it x[1]
    if (!exists("product_all")) {
      product_all <- x[1]
      # Otherwise take product_all times x[i] and store it in product_all
    } else {
      product_all <- product_all * x[i]
    }
  }
  
  # Second loop to populate y
  for (i in 1:length(x)) {
    # Each value of y is just product_all divided by the
    # ith value of x
    y[i] <- product_all / x[i]
  }
  
  # NOTE: Does not work if there are 0s in x
  # If there is one 0, we just need to replace one value
  if (!is.na(match(0, x)) & length(match(0, x) == 1)) {
    zero_pos <- match(0, x)
    
    x_aug <- x[-zero_pos]
    
    # Sets product_zero using loop
    for (j in 1:length(x_aug)) {
      if (!exists("product_x_aug")) {
        product_x_aug <- x_aug[1]
      } else {
        product_x_aug <- product_x_aug * x_aug[j]
      }
    }
    
    y[zero_pos] <- product_x_aug
  }
  
  # If there is more than one 0
  if (!is.na(match(0, x)) & length(match(0, x) > 1)) {
    # Create a vector y, the same length of x, filled with 0s
    y <- rep(0, length(x))
  }
  
  # Returns y
  y
}

```
    


## Exercise 2

Write a function f(x) to accept an integer vector, and returns a vector with those numbers … except for: 
   multiples of 3 = “Fizz” 
   multiples of 5 = “Buzz”. 
   multiples of 3 and 5 = “FizzBuzz”
   
### Example

    > x <- 1:20
    > f(1:20) 
    # 1 2 Fizz 4 Buzz Fizz 7 8 Fizz Buzz 11 Fizz 13 14 FizzBuzz 16 17 Fizz 19 Buzz


### Solution

```{r}
f <- function(x) {
  ifelse(x%%3 == 0
         & x%%5 == 0,
         "FizzBuzz",
  ifelse(x%%3 == 0,
         "Fizz",
  ifelse(x%%5 == 0,
         "Buzz",
         x)))
  
}
```
